Storm Surges

### 2014-07-15 RoughWork/manuscript.tex
Removed background in factors plots/discussion.

\begin{figure}[h]
\centering
\includegraphics[scale=0.5]{Figures/bathy.pdf}
\caption{Model domain (highlighted area) including bathymetry, rivers (blue circles), and storm surge locations of interest  (red stars).}\label{fig:domain}
\end{figure}
%image needs some work: colour? labels for JdF, SoG, cities...

@@

Next, an assessment of the factors that are most important in scenarios leading up to storm surges is presented in Figure \ref{fig:factors}. The modelled residual from three simulations have been plotted. First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly, is shown in blue. The modelled sea surface height has not been corrected as in the previous plot. Additionally, a simulation with all forcing except the atmospheric forcing is presented in green and a simulation with all forcing except the sea surface height anomaly is shown in red. It is clear that the sea surface height anomaly contributes most significantly to the surge at each of these locations. When the sea surface height anomaly is not included as a forcing condition, the surge drops from 78 cm to 28 cm at Point Atkinson. Additionally, the timing of the maximum surge occurs much later. A similar drop in the maximum surge and delay in the timing of the maximum surge is observed at the other locations.

The contribution to the surge due to atmospheric forcing appears to be geographically dependent, its removal leading to a drop (about 15 cm) in the maximum surge at Point Atkinson and Campbell River but a slight increase (about 6 cm) at Victoria and Patricia Bay. Additionally, at Point Atkinson the timing of the maximum surge is earlier by about two hours when atmospheric effects are neglected. 
%Worried that this might be due to pressure/mountains at PA and CR
%maybe look at winds to see why Vic/Pat Bay increased. Plan: map of wind vectors from CGRF (not for paper, just for me in aiding with explanation)

These comparisons point to the importance of including, as a forcing condition at the open bounday, the surge entering the domain from the Pacific Ocean. Storms travelling over the Pacific Ocean torwads the west coast of the northern United States and Canada can induce surges along the coastline which then enter the Salish Sea sytem through the Strait of Juan de Fuca. In a model domain that does not inlcude the Pacific Ocean, this effect must be inlcuded as a boundary condition. Previous simulations by \citet{murty1995storm} suggest that the inverese barometer effect and the surge entering the system through the Strait of Juan de Fuca contribute most significantly to storm surges in this domain.  A detailed assessment of the effect of the local atmospheric condtions and winds is included later where simulations have been performed with high resolution atmospheric forcing data.

\begin{figure}
\centering
\includegraphics[scale=0.6]{Figures/feb2006_factors.pdf}
\caption{Comparison of the modelled residual for three simulations: A simulation with all forcing (blue), a simulation without atmospheric forcing (green), and a simulation without the sea surface height anomaly (red). }
\label{fig:factors}
\end{figure}
%not sure about the colours, but different line styles is really hard to make out.

>>>>

\begin{figure}[h]
\centering
\includegraphics[scale=0.5]{Figures/bathy.pdf}
\caption{Model domain (highlighted area) including bathymetry, rivers (blue circles), and storm surge locations of interest  (red stars).}\label{fig:domain}
\end{figure}
%image needs some work: labels for tide comparisons?

@@

Next, an assessment of the factors that are most important in scenarios leading up to storm surges is presented in Figure \ref{fig:factors}. The modelled residual from three simulations have been plotted. First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly, is shown in blue. Additionally, a simulation with all forcing except the atmospheric forcing is presented in green and a simulation with all forcing except the sea surface height anomaly is shown in red. It is clear that the sea surface height anomaly contributes most significantly to the surge at each of these locations. When the sea surface height anomaly is not included as a forcing condition, the surge drops from 70 cm to 15.6 cm at Point Atkinson. Additionally, the timing of the maximum surge occurs much later. A similar drop in the maximum surge and delay in the timing of the maximum surge is observed at the other locations. Note that the press

The contribution to the surge due to atmospheric forcing appears to be geographically dependent, its removal leading to a drop of about 5 cm in the maximum surge at Point Atkinson and 10 cm at Campbell River. When atmospheric forces are neglected the timing of the maximum surge is earlier at Point Atkinson but later at Campbell River. Campbell River exhibits a strong response to atmopsheric conditions during midday on Feb 4, as demonstrated in a two sharp spikes in both the blue and red curves arounf 12:00 hr The green curve, which does not include any atmopsheric focring, does not exhibit this beahvious. The result is a peak surge earlier than in day than the no atmospheric forcing case. This response is also noted in the observations at Campbell River in Figure \ref{fig:feb2006} where the maxmium surge occurs earlier than at the other locations. At Victoria and Patricia Bay, the surge amplitude is mainly unaffected by the atmospheric forcing.  
%maybe look at winds/pressure to see why Vic/Pat Bay aren't affected?. Plan: map of wind vectors from CGRF (not for paper, just for me in aiding with explanation)

These comparisons point to the importance of including, as a forcing condition at the open bounday, the surge entering the domain from the Pacific Ocean. Storms travelling over the Pacific Ocean torwads the west coast of the northern United States and Canada can induce elevated sea levels along the coastline which then enter the Salish Sea sytem through the Strait of Juan de Fuca. In a model domain that does not inlcude the Pacific Ocean, this effect must be inlcuded as a boundary condition. Previous simulations by \citet{murty1995storm} suggest that the inverese barometer effect and the surge entering the system through the Strait of Juan de Fuca contribute most significantly to storm surges in this domain.  A detailed assessment of the effect of the local atmospheric condtions and winds is included later where simulations have been performed with high resolution atmospheric forcing data.

\begin{figure}
\centering
\includegraphics[scale=0.6]{Figures/feb2006_factors_bg.pdf}
\caption{Comparison of the modelled residual for three simulations: A simulation with all forcing (blue), a simulation without atmospheric forcing (green), and a simulation without the sea surface height anomaly (red). }
\label{fig:factors}
\end{figure}
%not sure about the colours, but different line styles is really hard to make out.
